Supporting Ecological Decision Making Using Feature-Selection and Variable Importance

### 2020-03-23 docs/00-manuscripts/ieee/ieee-manuscript.tex
write disc: feature selection

% learner
% - SVM Car is the overall winner
% - Ridge-CV hit and miss (hit = NRI, miss = VI)
% - all MBO penalized methods have no difference across tasks
SVM combined with the 'Carscore' filter achieved the best performance (RMSE of 27.98) (\autoref{tab:perf-top-10}).
%L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
Regression with Ridge penalty (L2) showed a high variance when comparing results across tasks: In two out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI, HR-VI, VI). For HR-NRI and HR the error was around 40 RMSE, and for task NRI the Ridge learner achieved almost the same score as the best performing model (32.9 RMSE) (\autoref{tab:perf-top-10}).
In all settings for which Ridge showed such a high error, only one observation in one fold was predicted which such a high value (in the millions).
This then caused the error estimates of these folds and the average estimate across all folds to be out of bounds.
% FIXME: All penalized methods tuned using MBO showed exactly the same error across all tasks.
% FIXME: Mention the bad perf of RIDGE on VI

@@

\subsubsection{Algorithm differences}
An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
% FIXME: Add refs
In the environmental community we have rarely seen a study in the last years in which SVM outperforms other models that clearly.
The absolute difference of around three percentage points is also worth noting: We could not find a study in which SVM came out as the best model in such clear way.
All of these statements are based on the assumption that a proper hyperparameter tuning was conducted for all models (which we also claim for ourselves in this study).

@@

% FIXME: Talk about L1 and L2 models

\subsubsection{Feature set differences}
One objective which this study aimed to answer was whether expert-based or data-driven feature engineering has a positive influence on model performance.
With respect to \autoref{fig:perf-result} one can say that overall the different feature sets did not result in substantial changes in performance.
RF and XGBoost showed a somewhat surprising decrease in performance of about six percentage points on task VI compared to all others.
We have no explanation for this specific result, especially because the SVM does not show this difference.
% FIXME: Talk about L1 and L2 models -> there we have quite some differences!

@@

% - Ensemble FS shows no advantage to simple methods -> maybe to less features, features too correlated? check diff between effect on NRI and VI
% - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
% - When calculating multiple filters, consider to only use ensemble filter?
% - PCA shows similar performance but probably requires tuning of main components -> check runtime!

@@

\begin{figure} [ht]
	\begin{center}
		\includegraphics[width=0.48\textwidth] {fi-hr-ale-1.pdf}
		\caption{ALE plots on dataset HR of SVM learner. Subset showing the ten most important features according to the permutation-based variable importance.}\label{fig:fi-hr-ale}
	\end{center}
\end{figure}

@@

\begin{figure} [ht]
	\begin{center}
		\includegraphics[width=0.48\textwidth] {fi-vi-ale-1.pdf}
		\caption{ALE plots on dataset VI of SVM learner. Subset showing the ten most important features according to the permutation-based variable importance.}\label{fig:fi-vi-ale}
	\end{center}
\end{figure}

>>>>

% learner
% - SVM Car is the overall winner
% - Ridge-CV hit and miss (hit = NRI, miss = VI)
% - all MBO penalized methods have no difference across tasks
SVM combined with the 'Carscore' filter achieved the best performance (RMSE of 27.98) (\autoref{tab:perf-top-10}).
%L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
Regression with Ridge penalty (L2) showed a high variance when comparing results across tasks: In two out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI, HR-VI, VI). For HR-NRI and HR the error was around 40 RMSE, and for task NRI the Ridge learner achieved almost the same score as the best performing model (32.9 RMSE) (\autoref{tab:perf-top-10}).
In all settings for which Ridge showed such a high error, only one observation in one fold was predicted which such a high value (in the millions).
This then caused the error estimates of these folds and the average estimate across all folds to be out of bounds.

FIXME: Mention: All penalized methods tuned using MBO showed exactly the same error across all tasks.
FIXME: Mention the bad perf of RIDGE on VI

@@

\subsubsection{Algorithm differences}
An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).

FIXME: Add refs
There were few environmental modeling studies only in recent years in which SVM outperformed other models that clearly.
The absolute difference of around three percentage points is also worth noting:
We could not find a study in which SVM came out as the best model in such clear way.
All of these statements are based on the assumption that a proper hyperparameter tuning was conducted for all models (which we also claim for ourselves in this study).

@@

FIXME: Talk about L1 and L2 model performance

\subsubsection{Feature set differences}
One objective which this study aimed to answer was whether expert-based or data-driven feature engineering has a positive influence on model performance.
With respect to \autoref{fig:perf-result} one can say that overall the different feature sets did not result in substantial changes in performance.
RF and XGBoost showed a somewhat surprising decrease in performance of about six percentage points on task VI compared to all others.
We have no explanation for this specific result, especially because the SVM does not show this difference.

FIXME: Talk about L1 and L2 models -> there we have quite some differences!

@@

% - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
The usefulness of filters in this study was apparent across all feature sets, even when $p$ was rather small (HR).
Even though the effect of feature selection methods varied across algorithms, small to substantial improvement in predictive performance were achieved in almost all cases.
No optimal algorithm/filter/dataset setting was observed in this study, leading to the conclusion that one needs to try out various combinations of such to find the best performing one.

% - Ensemble FS shows no advantage to simple methods -> maybe to less features, features too correlated? check diff between effect on NRI and VI
We have no explanation why the Borda ensemble method used in this study did not score better than the average performance of other filter methods.
Ensemble methods shine in complicated scenarios when simple filter methods face problems due to their focus on certain areas of features.
We assume that the datasets used in this study were not complex enough to bring up these problems for simple features.
Therefore it makes sense that, due to its averaging nature, the Borda filter scored average results across all instances.
However, this cannot be known upfront and we believe that an ensemble filter should be part of every filter portfolio when doing filter-based feature selection.

% - PCA shows similar performance but probably requires tuning of main components -> check runtime!
The PCA approach did neither show the best or worst result in the lineup.
It shines in terms of runtime since it reduces the amount of features highly.
However, it also removes the option for a post-analysis of feature importance by squashing predictors into principal components and users need to decided upfront what they value more.
Since filters are only calculated once due to caching, the runtime advantage might in fact be neglibile in practice.

@@

\begin{figure} [ht]
	\begin{center}
		\includegraphics[width=0.48\textwidth] {fi-hr-ale-1.pdf}
		\caption{ALE plots on dataset HR of SVM learner. Subset showing the ten most important features according to the permutation-based variable importance. The y-axis shows the deviation to the mean prediction for each feature, with the mean prediction being centered at zero.}\label{fig:fi-hr-ale}
	\end{center}
\end{figure}

@@

\begin{figure} [ht]
	\begin{center}
		\includegraphics[width=0.48\textwidth] {fi-vi-ale-1.pdf}
		\caption{ALE plots on dataset VI of SVM learner. Subset showing the ten most important features according to the permutation-based variable importance. The y-axis shows the deviation to the mean prediction for each feature, with the mean prediction being centered at zero.}\label{fig:fi-vi-ale}
	\end{center}
\end{figure}
